A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-20 paper/paper.Rnw
ROC + examples comments and changes

The full model has an AUC of \Sexpr{round(whole_model_roc_df$auc,2)}, and the AUC for individual classes ranges from \Sexpr{round(min(aucs$data$auc),2)} (for \Sexpr{sprintf("%s",classes[which.min(aucs$data$auc)])}) to \Sexpr{round(max(aucs$data$auc),2)}  (for \Sexpr{sprintf("%s",classes[which.max(aucs$data$auc)])}).
While the class performances do vary slightly, each ROC curve is the same general shape and performs significantly better than random chance. The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. We will use the class-wise \mt{classwise?} EER to set a threshold corresponding to each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.

\paragraph{Model Predictions---Examples}
The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective \mt{model-assigned} nine class predictions. \autoref{fig:exemplars} and \autoref{fig:circle-text-consistency} show model predictions for a selection of images. \mt{Each image shown in the following examples belongs to either the validation or test data, so none of the images were used during model training. In these figures, the values in each row are the predicted probabilities for the nine classes in the corresponding image image, such that each column represents predictions for a given class. The saturation of blue color corresponds to the strength of the predicted probability, and the navy blue border around an image indicates that the predicted probability is above the EER threshold for the given class. Note that the EER is not the same for all classes; in other words, the same probablity value assigned across different categories may yield different conclusions about whether the model predicts the shape is present or absent in the image.}

The first rows of \autoref{fig:exemplars} are images which contain only one of the nine classes. Most of the predictions for these images are \mt{promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above EER. The next image belongs to the ``other" category, as it does not contain any of the nine classes; the model correctly assigns near-zero probabilities to each class. The final rows are of images that belong to more than one class category. Again, predictions appear quite good, with most of the prominent shapes correctly identified above EER. \fix{These are mostly new examples from our last meeting because many of the previous images were in the training set... Grrrrr..... I have two or three too many, not sure about color contrast within images, etc., feel free to delete a few or we can discuss them.} [Expand on specific misclassifications more, after determining which examples to delete.] A later section will discuss possible causes for these class confusions.}

<<eers, include = F>>=
eers <- purrr::map_dbl(aucs$data$eer, function(x){x$thresholds}) %>%
  setNames(classes)
@

<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T>>=
single_exemplars <- list.files(file.path(wd, "images", "consistency",
                                         "single-exemplar"), full.names = T)
multi_exemplars <- list.files(file.path(wd, "images", "consistency",
                                        "multi-exemplar"), full.names = T)
pred_prob_plot(rev(c(single_exemplars, multi_exemplars)),
               loaded_model, sort = F, eer = eers)
@

\mt{While figure \ref{fig:exemplars} shows representative examples of classifications across multiple shape categories, it is also useful to examine many cases within a given shape to get a sense of model consistency.} \mt{\autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text. The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text. The remaining images contain only one of either text or circles, as well as other possible shape categories. In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final \fix{N} images contain pure circles that do not appear to belong to text. The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories. This confusion is likely due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text. In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape. More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.}

\mt{Side note: I keep futzing with this section. I am going to forcibly move on and welcome Susan's feedback, and not allow myself to change any more of the consistency plots.}

@@

\mt{\paragraph{Confusion Matrix}}
\mt{Figures \ref{fig:exemplars} and \ref{fig:circle-text-consistency} show predictions for a number of individual images. Many of the misclassifications in these examples, however, are not just seen in individual images, but rather exist systematically between certain classes.} A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified. \mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}. %mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column. Thus, }
The values along the diagonal \mt{of the matrix} represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions. For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels. \mt{As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise). Note that this matrix is not symmetric---the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.}

>>>>

The full model has an AUC of \Sexpr{round(whole_model_roc_df$auc,2)}, and the AUC for individual classes ranges from \Sexpr{round(min(aucs$data$auc),2)} (for \Sexpr{sprintf("%s",classes[which.min(aucs$data$auc)])}) to \Sexpr{round(max(aucs$data$auc),2)}  (for \Sexpr{sprintf("%s",classes[which.max(aucs$data$auc)])}).
While the class performances do vary slightly, each ROC curve \svp{indicates that the model} performs significantly better than random chance. The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. We will use \svp{a class specific} EER to set \svp{an individual} threshold \svp{for} each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.

\paragraph{Model Predictions---Examples}
The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with \svp{the nine class predictions assigned by the model}. \autoref{fig:exemplars} and \autoref{fig:circle-text-consistency} show model predictions for a selection of images. Each image shown in the following examples belongs to either the validation or test data, so none of the images were used during model training. In these figures, the values in each row are the predicted probabilities for the nine classes in the corresponding image image, such that each column represents predictions for a given class. The saturation of blue color corresponds to the strength of the predicted probability, and the navy blue border around an image indicates that the predicted probability is above the EER threshold for the given class. Note that the EER is not the same for all classes; in other words, the same probablity value assigned across different categories may yield different conclusions about whether the model predicts the shape is present or absent in the image.

The first rows of \autoref{fig:exemplars} are images which contain only one of the nine classes. Most of the predictions for these images are promising; there are some false-positive labels that indicate confusions between circle/text and polygon/quad, but all of the true shapes are corectly identified by the model above EER. The next image belongs to the ``other" category, as it does not contain any of the nine classes; the model correctly assigns near-zero probabilities to each class. The final rows are of images that belong to more than one class category. Again, predictions \svp{are largely accurate}, with most of the prominent shapes correctly identified above EER. \fix{These are mostly new examples from our last meeting because many of the previous images were in the training set... Grrrrr..... I have two or three too many, not sure about color contrast within images, etc., feel free to delete a few or we can discuss them. [Expand on specific misclassifications more, after determining which examples to delete.]}
\fix{We need to make sure that the x axis labels aren't crammed together... Overall the examples are good, but I think I might get rid of the all-white DC image, and possibly the one right below it as well. We can discuss on Monday.}
<<eers, include = F>>=
eers <- purrr::map_dbl(aucs$data$eer, function(x){x$thresholds}) %>%
  setNames(classes)
@

<<exemplars, fig.cap = "Examples containing a single shape", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 8, fig.height = 12>>=
single_exemplars <- list.files(file.path(wd, "images", "consistency",
                                         "single-exemplar"), full.names = T)
multi_exemplars <- list.files(file.path(wd, "images", "consistency",
                                        "multi-exemplar"), full.names = T)
pred_prob_plot(rev(c(single_exemplars, multi_exemplars)),
               loaded_model, sort = F, eer = eers)
@

While figure \ref{fig:exemplars} shows representative examples of classifications across multiple shape categories, it is also useful to examine many cases within a given shape to get a sense of model consistency. \autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text. \fix{this isn't accurate to what I'm seeing...} \fix{The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text. The remaining images contain only one of either text or circles, as well as other possible shape categories.} In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final \fix{N} images contain pure circles that do not appear to belong to text. The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having dificulty distinguishing the two categories. \svp{While the circle predictions are above the EER for letters such as C and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular.} This confusion is \svp{complicated} due to the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text. In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape. More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.

% \mt{Side note: I keep futzing with this section. I am going to forcibly move on and welcome Susan's feedback, and not allow myself to change any more of the consistency plots.} \svp{Good strategy. Biggest feedback is to change the text to actually match the image as it is currently, but I get why you stopped and I'm proud of you for noticing that it was a problem and finding a solution. :) Nice job.}

@@

\paragraph{Confusion Matrix}
Figures \ref{fig:exemplars} and \ref{fig:circle-text-consistency} show predictions for a number of individual images. Many of the misclassifications in these examples, however, are not just seen in individual images, but rather exist systematically between certain classes. % Nicely phrased!!! I couldn't have done it better myself.
A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model across all test data, and is a helpful diagnostic tool to understand which classes are systematically misclassified. The confusion matrix for the fitted model is presented in \autoref{fig:confusion_matrix.} %mt{Each value in the matrix represents the probability of assigning the label associated with that value's corresponding row, conditional upon the image truly belonging to the category associated with the value's column. Thus, }
The values along the diagonal of the matrix represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions. \svp{Thus, this might be considered a conditional confusion matrix, because we condition the off diagonal probabilities on the label not being correct.} For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any images that truly contains both triangles and quadrilaterals are removed before calculating the proportion of false quadrilateral labels. As a result of this adjustment, the off-diagonal values of the confusion matrix are the proportion of images within a true class (column-wise) that are falsely labeled as containing another class that has not been labeled in the image (row-wise). Note that this matrix is not symmetric: the proportion of images containing text that are falsely predicted to contain circles is not necessarily equal to the proportion of images containing circles that are falsely predicted to contain text.
